---
title: "Bivariate Inference_Week7"
author: "Michael Cao"
date: 2/24/25
format:
  html:
    self-contained: true
editor: source
execute:
  echo: false
  message: false
  warning: false
  cache: true 
---
```{r}
library(tidyverse)
library(gssr)
```

```{r}
gss2022 <- gss_get_yr(year = 2022) |>
  haven::zap_labels()
```

**Hypothesis Test #1**

For the binary variable *pres20* from the 2022 GSS ("Did you vote for Biden"), grouped the variable *sex* (M/F). 

```{r, results = "hide"}
gss2022$pres20 <- ifelse(gss2022$pres16 == 1, 1, 0) 
sum(!is.na(gss2022$pres20))
gss2022$sex <- ifelse(gss2022$sex == 1, 1, 0)
sum(!is.na(gss2022$sex))
```

Let's ask: is there a difference in votes for Joe Biden between men and women? 

$H_{0} : p_{male} - p_{female} = 0$

$H_{1} : p_{male} - p_{female} \neq 0$

For $\alpha = 0.01$

```{r}
p1_pressex <- mean(gss2022$pres20[gss2022$sex == "1"], na.rm = TRUE)
p2_pressex <- mean(gss2022$pres20[gss2022$sex == "0"], na.rm = TRUE)
phat_total_pressex <- mean(gss2022$pres20, na.rm = TRUE)
n1_pressex <- sum(!is.na(gss2022$pres20[gss2022$sex == "1"]))
n2_pressex <- sum(!is.na(gss2022$pres20[gss2022$sex == "0"]))
```

$\hat{p}_{male} = `r round(p1_pressex, 3)`$

$\hat{p}_{female} = `r round(p2_pressex, 3)`$

$\hat{p}_{male - female} = `r round(p1_pressex - p2_pressex, 3)`$

```{r}
data_frame1 <- gss2022 |>
  select(pres20, sex) |>
  drop_na() |>
  mutate(pres20 = if_else(pres20 == 1, 1, 0),
         sex = if_else(sex == 1, "Male", "Female")) |>
           group_by(sex) |>
           summarize(prop = mean(pres20)) |>
ggplot(aes(x = sex,
           y = prop * 100,
           fill = sex)) + 
  geom_col() + 
  scale_fill_manual(values = c("Male" = "#057f8d", "Female" = "#ff7261")) +
  labs(title = "Voting for Biden by Sex",
       x = "Sex",
       y = "% Voters", 
       caption = "Source: 2022 GSS") +
  theme_classic()
print(data_frame1)
```

```{r, echo = TRUE}
z_stat1 <- abs((p1_pressex - p2_pressex) / sqrt(phat_total_pressex * (1 - phat_total_pressex) * (1 / n1_pressex + 1 / n2_pressex)))
round(z_stat1, 3) # Calculating the z-statistic
p_stat1 <- 2 * (1 - pnorm(abs(z_stat1)))
p_stat1 # Calculating the p-value
```

```{r}
obsdiff_test1 <- p1_pressex - p2_pressex
SE_test1 <- sqrt(p1_pressex * (1 - p1_pressex) / n1_pressex + p2_pressex * (1 - p2_pressex) / n2_pressex)
lower99_test1 <- obsdiff_test1 - (2.57 * SE_test1)
upper99_test1 <- obsdiff_test1 + (2.57 * SE_test1)
```

Calculating a two-tailed p-value, our result is extremely small, which means we can reject the null hypothesis that there is no difference in men and women who voted Biden. It is beyond extreme that we would obtain our sample proportion difference given the null distribution. Translating this to a confidence interval, the 99% confidence interval for this sample would be $\{`r round(lower99_test1, 3)`, `r round(upper99_test1, 4)`\}$. This interval does not include 0, so alternatively, we can use it to reject the null hypothesis. 

**Hypothesis Test #2** 

For the variable *life* ("Do you find life exciting or not") from the 2022 GSS, grouped by *age* (Younger/Older).

```{r, results = "hide"}
gss2022$life <- ifelse(gss2022$pres16 == 1, 1, 0) 
sum(!is.na(gss2022$life))
gss2022$age <- ifelse(gss2022$age <= 50, 1, 0)
sum(!is.na(gss2022$age))
```

Let's ask: is there a difference in life excitement between the younger and the older? 

$H_{0} : p_{young} - p_{old} = 0$

$H_{1} : p_{young} - p_{old} \neq 0$

For $\alpha = 0.01$

```{r}
p1_lifeage <- mean(gss2022$life[gss2022$age == "1"], na.rm = TRUE)
p2_lifeage <- mean(gss2022$life[gss2022$age == "0"], na.rm = TRUE)
phat_total_lifeage <- mean(gss2022$life, na.rm = TRUE)
n1_lifeage <- sum(!is.na(gss2022$life[gss2022$age == "1"]))
n2_lifeage <- sum(!is.na(gss2022$life[gss2022$age == "0"]))
```

$\hat{p}_{male} = `r round(p1_lifeage, 3)`$

$\hat{p}_{female} = `r round(p2_lifeage, 3)`$

$\hat{p}_{male - female} = `r round(p1_lifeage - p2_lifeage, 3)`$

```{r}
data_frame2 <- gss2022 |>
  select(life, age) |>
  drop_na() |>
  mutate(life = if_else(life == 1, 1, 0),
         age = if_else(age == 1, "Younger", "Older")) |>
           group_by(age) |>
           summarize(prop = mean(life)) |>
ggplot(aes(x = age,
           y = prop * 100,
           fill = age)) + 
  geom_col() + 
  scale_fill_manual(values = c("Younger" = "#057f8d", "Older" = "#ff7261")) +
  labs(title = "Exciting Lives by Age",
       x = "Age (Older = 50+)",
       y = "% with Exciting Lives", 
       caption = "Source: 2022 GSS") +
  theme_classic()
print(data_frame2)
```

```{r, echo = TRUE}
z_stat2 <- abs((p1_lifeage - p2_lifeage) / sqrt(phat_total_lifeage * (1 - phat_total_lifeage) * (1 / n1_lifeage + 1 / n2_lifeage)))
round(z_stat2, 3) # Calculating the z-statistic
p_stat2 <- 2 * (1 - pnorm(abs(z_stat2)))
p_stat2 # Calculating the p-value
```

```{r}
obsdiff_test2 <- p1_lifeage - p2_lifeage
SE_test2 <- sqrt(p1_lifeage * (1 - p1_lifeage) / n1_lifeage + p2_lifeage * (1 - p2_lifeage) / n2_lifeage)
lower99_test2 <- obsdiff_test2 - (2.57 * SE_test2)
upper99_test2 <- obsdiff_test2 + (2.57 * SE_test2)
```

Calculating a two-tailed p-value, our result is pretty small, which means we fail to reject the null hypothesis that there is no difference in life excitement among the younger and older. It pretty unlikely that we would obtain our sample proportion difference given the null distribution and our conservative alpha level. Translating this to a confidence interval, the 99% confidence interval for this sample would be $\{`r round(lower99_test2, 5)`, `r round(upper99_test2, 3)`\}$. This interval does not include 0, so alternatively, we can use it to reject the null hypothesis. 

**Hypothesis Test #3** 

For the variable *nmbrkids* ("How many children have you ever had") grouped by *degree* status (at least a BA/no BA).

```{r, results = "hide"}
sum(!is.na(gss2022$nmbrkids))
gss2022$educ <- ifelse(gss2022$degree >=3, 1, 0)
sum(!is.na(gss2022$degree))
```

Let's ask: is there a difference in the number of children between those with and without a BA degree? 

$H_{0} : \mu_{BA+} - \mu_{<BA} = 0$

$H_{1} : \mu_{BA+} - p_{<BA} \neq 0$

For $\alpha = 0.01$

```{r}
mean1_childcollege <- mean(gss2022$nmbrkids[gss2022$degree == "1"], na.rm = TRUE)
mean2_childcollege <- mean(gss2022$nmbrkids[gss2022$degree == "0"], na.rm = TRUE)
sd1_childcollege <- sd(gss2022$nmbrkids[gss2022$degree == "1"], na.rm = TRUE)
sd2_childcollege <- sd(gss2022$nmbrkids[gss2022$degree == "0"], na.rm = TRUE)
n1_childcollege <- sum(!is.na(gss2022$nmbrkids[gss2022$degree == "1"]))
n2_childcollege <- sum(!is.na(gss2022$nmbrkids[gss2022$degree == "0"]))
se_diff_childcollege <- sqrt((sd1_childcollege / n1_childcollege) + (sd2_childcollege / n2_childcollege))
```

$\bar{x}_{BA_+} = `r round(mean1_childcollege, 3)`$

$\bar{x}_{<BA} = `r round(mean2_childcollege, 3)`$

$\bar{x}_{BA_+ - <BA} = `r round(mean1_childcollege - mean2_childcollege, 3)`$

```{r}
data_frame3 <- gss2022 |>
  select(nmbrkids, degree) |>
  drop_na() |>
  mutate(degree = if_else(degree >= 3, "BA+", "<BA")) |>
           group_by(degree) |>
           summarize(mkids = mean(nmbrkids)) |>
ggplot(aes(x = degree,
           y = mkids,
           fill = degree)) + 
  geom_col() +
  scale_fill_manual(values = c("BA+" = "#057f8d", "<BA" = "#ff7261")) +
  labs(title = "Number of Kids by Degree",
       x = "Degree",
       y = "Mean Number of Kids", 
       caption = "Source: 2022 GSS") +
  theme_classic()
data_frame3
```

```{r}
data_frame3_distribution <- gss2022 |>
  select(nmbrkids, degree) |>
  drop_na() |>
  mutate(degree = if_else(degree >= 3, "BA+", "<BA")) |>
  ggplot(aes(x = nmbrkids,
        fill = degree)) +
        geom_density(alpha = 0.5,
                     color = NA) +
        scale_fill_manual(values = c("BA+" = "#057f8d", "<BA" = "#ff7261")) +
        xlim(0,10) +
        labs(title = "Distribution of Number of Kids by Degree",
            x = "nmbrkids",
            caption = "2022 GSS") 
data_frame3_distribution
```

```{r, echo = TRUE}
z_stat3 <- abs((mean1_childcollege - mean2_childcollege) / se_diff_childcollege)
round(z_stat3, 3) # Calculating the z-statistic
p_stat3 <- 2 * (1 - pnorm(abs(z_stat3)))
p_stat3 # Calculating the p-value
```

```{r}
obsdiff_test3 <- mean1_childcollege - mean2_childcollege
lower99_test3 <- obsdiff_test3 - (2.57 * se_diff_childcollege)
upper99_test3 <- obsdiff_test3 + (2.57 * se_diff_childcollege)
```

Calculating a two-tailed p-value, our result is very small, which means we can reject the null hypothesis that there is no difference in the number of children  among those with at least a BA degree and those without. It is highly unlikely that we would obtain our sample proportion difference given the null distribution. Translating this to a confidence interval, the 99% confidence interval for this sample would be $\{`r round(lower99_test3, 3)`, `r round(upper99_test3, 3)`\}$. This interval does not include 0, so alternatively, we can use it to reject the null hypothesis. 

**Hypothesis Test #4** 

For the variable *weight* grouped by *divorce* status (Divorced/Non-Divorced). 

```{r, results = "hide"}
sum(!is.na(gss2022$weight))
gss2022$divorce <- ifelse(gss2022$divorce == 1, 1, 0)
sum(!is.na(gss2022$divorce))
```

Let's ask: is there a difference in weight between the divorced and non-divorced? 

$H_0 : \mu_{divorced} - \mu_{non-divorced} = 0$

$H_0 : \mu_{divorced} - \mu_{non-divorced} \neq 0$

For $\alpha = 0.01$

```{r}
mean1_weightdiv <- mean(gss2022$weight[gss2022$divorce == "1"], na.rm = TRUE)
mean2_weightdiv <- mean(gss2022$weight[gss2022$divorce == "0"], na.rm = TRUE)
sd1_weightdiv <- sd(gss2022$weight[gss2022$divorce == "1"], na.rm = TRUE)
sd2_weightdiv <- sd(gss2022$weight[gss2022$divorce == "0"], na.rm = TRUE)
n1_weightdiv <- sum(!is.na(gss2022$weight[gss2022$divorce == "1"]))
n2_weightdiv <- sum(!is.na(gss2022$weight[gss2022$divorce == "0"]))
se_diff_weightdiv <- sqrt((sd1_weightdiv / n1_weightdiv) + (sd2_weightdiv / n2_weightdiv))
```

$\bar{x}_{divorced} = `r round(mean1_weightdiv, 3)`$

$\bar{x}_{non-divorced} = `r round(mean2_weightdiv, 3)`$

$\bar{x}_{divorced - non-divorced} = `r round(mean1_weightdiv - mean2_weightdiv, 3)`$

```{r}
data_frame4 <- gss2022 |>
  select(weight, divorce) |>
  drop_na() |>
  mutate(divorce = if_else(divorce == 1, "Divorced", "Non-Divorced")) |>
           group_by(divorce) |>
           summarize(mweight = mean(weight)) |>
ggplot(aes(x = divorce,
           y = mweight,
           fill = divorce)) + 
  geom_col() +
  scale_fill_manual(values = c("Divorced" = "#057f8d", "Non-Divorced" = "#ff7261")) +
  labs(title = "Weight by Divorce Status",
       x = "Divorce Status",
       y = "Mean Weight (lbs)", 
       caption = "Source: 2022 GSS") +
  theme_classic()
data_frame4
```

```{r}
data_frame4_distribution <- gss2022 |>
  select(weight, divorce) |>
  drop_na() |>
  mutate(divorce = if_else(divorce == 1, "Divorced", "Non-Divorced")) |>
  ggplot(aes(x = weight,
        fill = divorce)) +
        geom_density(alpha = 0.3,
                     color = NA) +
        scale_fill_manual(values = c("Divorced" = "#057f8d", "Non-Divorced" = "#ff7261")) +
        xlim(0, 400) +
        labs(title = "Distribution of Weight by Degree Status",
            x = "weight (lbs)",
            caption = "2022 GSS") 
data_frame4_distribution
```

```{r, echo = TRUE}
z_stat4 <- abs((mean1_weightdiv - mean2_weightdiv) / se_diff_weightdiv)
round(z_stat4, 3) # Calculating the z-statistic
p_stat4 <- 2 * (1 - pnorm(abs(z_stat4)))
p_stat4 # Calculating the p-value
```

```{r}
obsdiff_test4 <- mean1_weightdiv - mean2_weightdiv
lower99_test4 <- obsdiff_test4 - (2.57 * se_diff_weightdiv)
upper99_test4 <- obsdiff_test4 + (2.57 * se_diff_weightdiv)
```

Calculating a two-tailed p-value, our result is basically 0, which means we can reject the null hypothesis that there is no difference in weight between the divorced and non-divorced. It is near impossible that we would obtain our sample proportion difference given the null distribution. Translating this to a confidence interval, the 99% confidence interval for this sample would be $\{`r round(lower99_test4, 3)`, `r round(upper99_test4, 3)`\}$. This interval does not include 0, so alternatively, we can use it to reject the null hypothesis. 

