---
title: "Univariate Stats_Week4"
author: "Michael Cao"
date: 2/5/25
format:
  html:
    self-contained: true
editor: visual
execute:
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(gssr)
```

```{r, results = "hide"}
gss2022 <- gss_get_yr(year = 2022) |>
  haven::zap_labels()
glimpse(gss2022$age)
mean(gss2022$age, na.rm = TRUE)
sd(gss2022$age, na.rm = TRUE)
```

**Exercise #1:**

Describing the center and spread for the variable *age*, drawn from the 2022 GSS.

```{r}
var1_plot <- gss2022 |>
  select(age) |>
  drop_na()

ggplot(var1_plot, aes(x = age)) +
  geom_histogram(fill = "lightgray", color = "black", binwidth = 7.1) +
  scale_x_continuous(breaks = seq(min(var1_plot$age), max(var1_plot$age), by = 7.1)) +
  labs(x = "Age (in years)",
       y = "Count",
       caption = "Source: 2022 GSS",
       title = "Distribution of age (n = 3893)") +
  theme_classic()
```

Mean ($\bar{x}_{age}$):

$\bar{x}_{age} = \frac{1}{n} \sum_{i=1}^{n} x_i$

$\bar{x}_{age} = `r mean(gss2022$age, na.rm = TRUE)`$

Standard Deviation ($s_{age}$):

$s_{age} = \sqrt{\frac{1}{n-1} \sum_{i=1}^{n}(x_i-\bar{x}_{age})^2}$

$s_{age} = `r sd(gss2022$age, na.rm = TRUE)`$

$\bar{x}_{age}$ measures the center of the *age* sample distribution, which here is balanced around 48.2 years old---so I'd say most of the sample tends to be middle-aged. $s_{age}$ tells us each observation (respondent's age) in the sample deviates from the mean (48.2 years) by a typical distance of 17.7 years. This seems pretty high, which would suggest a good amount of variability in sample ages, including both younger and older ages. 

Plotting the probability density distribution of *age*, and then approximating the normal distribution of *age*, given $\text{age} \sim N(`r round(mean(gss2022$age, na.rm = TRUE), 1)`, `r round(sd(gss2022$age, na.rm = TRUE), 1)`)$: 

```{r}
ggplot(var1_plot, aes(x = age)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightgray", color = "black", binwidth = 7.1) +
  geom_density(col = "#17EFF6", lwd = 1) + 
  stat_function(fun = dnorm, args = list(mean = mean(var1_plot$age), #drew this from the slides
                                        sd = sd(var1_plot$age)),
                col = "#F77A76", lwd = 1, linetype = "dashed") +  
  scale_x_continuous(breaks = seq(min(var1_plot$age), max(var1_plot$age), by = 7.1)) +
  labs(x = "Age (in years)",
       y = "Density",
       caption = "Source: 2022 GSS",
       title = "Probability Density Distribution of age Approximated as a Normal Distribution") +
  theme_classic()
```

```{r, results = "hide"}
glimpse(gss2022$hompop)
mean(gss2022$hompop, na.rm = TRUE)
sd(gss2022$hompop, na.rm = TRUE)
```

Comparing the normal approximation to the observed empirical distribution, we notice: 

* The normal approximation is simultaneously under-approximating the frequency of tail-end younger ages and older ages, while over-approximating the frequency of middle ages in the sample.

* As a consequence of this, the normal approximation seems to miss the slight bimodality of the empirical sample distribution that has peaks towards young and older ages.

* The normal approximation also assumes the sample distribution stretches to non-existent ages lesser or greater, respectively, than the minimum and maximum values of the sample. 

**Exercise #2:**

Describe the center and spread for the variable *hompop* (household size), drawn from the 2022 GSS. I use `geom_bar()` in my `ggplot()` argument because I treat *hompop* as discrete. 

```{r}
var2_plot <- gss2022 |> 
  select(hompop) |>
  drop_na()

ggplot(var2_plot, aes(x = hompop)) +
  geom_bar(fill = "lightgray", color = "black") +  # I use geom_bar in my ggplot argument because I treat *hompop* as discrete. 
  scale_x_continuous(breaks = 1:12) +
  labs(x = "Household Size", 
       y = "Count",
       caption = "Source: 2022 GSS",
       title = "Distribution of hompop (n = 2079)") +
  theme_classic()
```

Mean ($\bar{x}_{hompop}$):

$\bar{x}_{hompop} = \frac{1}{n} \sum_{i=1}^{n} x_i$

$\bar{x}_{hompop} = `r mean(gss2022$hompop, na.rm = TRUE)`$

Standard Deviation ($s_{hompop}$):

$s_{hompop} = \sqrt{\frac{1}{n-1} \sum_{i=1}^{n}(x_i-\bar{x}_{hompop})^2}$

$s_{hompop} = `r sd(gss2022$hompop, na.rm = TRUE)`$

$\bar{x}_{hompop}$ measures the center of the *hompop* sample distribution, which here is balanced around a household size of 2.4. I am not sure about exact relationships, but it appears the sample tends to live with another person (likely a partner or significant other). $s_{hompop}$ tells us each observation (respondent's household size) in the sample deviates from the mean (2.4) by a typical distance of 1.5. This suggests pretty low variability in sample household sizes, which indicates most household sizes don't get too smaller or bigger than 2.4. 

Plotting the probability density distribution of *hompop*, and then approximating the normal distribution of *hompop*, given $\text{hompop} \sim N(`r round(mean(gss2022$hompop, na.rm = TRUE), 1)`, `r round(sd(gss2022$hompop, na.rm = TRUE), 1)`)$:

```{r}
ggplot(var2_plot, aes(x = hompop)) +
  geom_histogram(aes(y = after_stat(density)), fill = "lightgray", color = "black", binwidth = 1) + 
  stat_function(fun = dnorm, #drew this from the slides
                args = list(mean = mean(var2_plot$hompop),
                            sd = sd(var2_plot$hompop)), 
                col = "#F77A76", lwd = 1, linetype = "dashed") + 
  geom_density(col = "#17EFF6", lwd = 1) + 
  scale_x_continuous(breaks = 1:12) +
  labs(x = "Household Size",
       y = "Density", 
       caption = "Source: 2022 GSS",
       title = "Probability Density Distribution of hompop Approximated as a Normal Distribution") + 
  theme_classic()
```

Comparing the normal approximation to the observed empirical distribution, we notice: 

* Though both distributions are unimodal, the normal approximation appears to miss the right-skew of the sample distribution. 

* Because the normal approximation assumes a symmetrical shape, a glaring issue is that it assumes negative household sizes, which is not possible. I suppose 12+ household sizes is theoretically possible, but extremely rare, which actually I think is captured by the normal approximation. 

* The normal approximation is also under-approximating the frequency of smaller household sizes in the sample distribution. 