---
title: "GLM_practice_Week11"
author: "Michael Cao"
date: 3/27/25
format:
  html:
    self-contained: true
editor: source
execute:
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(gssr)
```

```{r}
gss2022 <- gss_get_yr(year = 2022) |>
  haven::zap_labels()
```

**Models 1, 2, 3: Estimating Binary GLMs**

$biden_{i = 1} = \alpha + \beta(gunlaw_{i})$

predictor: favor stronger gun laws 

outcome: voted for Biden 

```{r}
data1 <- gss2022 |> 
  select(pres20, gunlaw) |>
  drop_na() |>
  mutate(biden = if_else(pres20 == 1, 1, 0),
         gunlaw = if_else(gunlaw == 1, 1, 0)) |>
  select(biden, gunlaw)
```

**Linear Probability Model**

```{r, echo = TRUE}
model1 <- glm(biden ~ gunlaw,
              data = data1,
              family = binomial(link = "identity")) 
 
broom::tidy(model1,
            conf.int = TRUE,
            conf.level = 0.99)
```

* The estimated probability of someone voting Biden given that they oppose stronger gun laws is 32.8% and 72.7% if they favor stronger gun laws.
* Someone who favors stronger gun laws is 39.9 percentage points more likely to have voted for Biden than someone who opposes stronger gun laws. 

**Calculating Log-Likelihood Manually and by Function**

```{r, echo = TRUE}
data1 <- 
  data1 |> 
  mutate(predprob = predict(model1,
                            type = "response")) |>
  mutate(l = dbinom(biden, 1, predprob),
         ll = log(l))
```

```{r, echo = TRUE, results = TRUE}
sum(data1$ll)
logLik(model1)
```

**Log Probability Model** 

```{r, echo = TRUE}
model2 <- glm(biden ~ gunlaw,
              data = data1,
              family = binomial(link = "log")) 
 
broom::tidy(model2,
            conf.int = TRUE,
            conf.level = 0.99)
```

```{r, echo = TRUE, results = TRUE}
exp(0.797)
```

* Someone who favors stronger gun laws has an estimated probability of voting Biden that is 2.21 times the estimated probability for someone is opposed to stronger gun laws. 

**Logit (Log Odds) Model** 

```{r}
model3 <- glm(biden ~ gunlaw,
              data = data1,
              family = binomial(link = "logit")) 
 
broom::tidy(model3,
            conf.int = TRUE,
            conf.level = 0.99)
```
```{r, echo = TRUE, results = TRUE}
exp(1.70)
```

* Someone who favors stronger gun laws has much high odds (5.5 times) of supporting Biden than the odds of someone who opposes stronger gun laws. 

**Gaussian Regression (Estimating a Linear Model)**

$weight_{i} = \alpha + \beta(evermarried_{i})$

predictor: ever-married 

outcome: weight 

```{r}
data2 <- gss2022 |>
  select(weight, marital) |>
  drop_na() |>
  mutate(evermarried = if_else(marital <= 4, 1, 0)) |>
  select(weight, evermarried)
```

```{r, echo = TRUE}
model4 <- lm(weight ~ evermarried,
             data = data2)
 
broom::tidy(model4,
            conf.int = TRUE,
            conf.level = 0.99)
```
* The estimated mean weight of someone who has never been married is 185 pounds. 
* Someone who has ever been married is estimated to 0.313 pounds more than someone who has never been married.

**Calculating Log-Likelihood Manually and by Function**

```{r, echo = TRUE}
data2 <- 
  data2 |>
    mutate(mean = predict(model4),
           sd = sd(model4$residuals)) |>
    mutate(l  = dnorm(weight,      
                      mean = mean,
                      sd = sd),
           ll = dnorm(weight,      
                      mean = mean,
                      sd = sd,
                      log = TRUE))
```

```{r, echo = TRUE, results = TRUE}
sum(data2$ll)
logLik(model4)
```

**Poisson Regression (Estimating a Count Model)**

$kids_{i} = \alpha + \beta(evermarried_{i})$

predictor = ever-married

outcome = number of kids 

```{r}
data3 <- gss2022 |>
  select(nmbrkids, marital) |>
  drop_na() |>
  mutate(kids = nmbrkids,
         evermarried = if_else(marital <= 4, 1, 0)) |>
  select(kids, evermarried)
```

```{r, echo = TRUE}
model5 <- glm(kids ~ evermarried,
             data = data3,
             family = poisson(link = "log"))
 
broom::tidy(model5,
            conf.int = TRUE,
            conf.level = 0.99)
```
```{r, echo = TRUE, results = TRUE}
exp(0.337)
exp(0.532)
```

* The estimated mean number of kids of someone who has never been married is 1.40. 
* Someone who has ever been married is estimated to have 1.70 times as many kids as someone who has never been married. 

**Calculating Log-Likelihood Manually and by Function**

```{r, echo = TRUE}
data3 <- 
  data3 |>
    mutate(mean = predict(model5, type = "response")) |>
    mutate(l  = dpois(kids,      
                      lambda = mean),
           ll = dpois(kids,      
                      lambda = mean,
                      log = TRUE))
```

```{r, echo = TRUE, results = TRUE}
sum(data3$ll)
logLik(model5)
```

