---
title: "Probability Exercise for SOC 722"
author: "Michael Cao"
date: 1/29/25
format:
  html:
    self-contained: true
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
#| echo: false
library(tidyverse)
library(tinytable)
library(gssr)
```

# Table #1

For the first 2x2 contingency table, I pulled and recoded the variables *polviews* (respondent's political views) and *cappun* (respondent's support for capital punishment) from the 2022 GSS.

```{r}
#| echo: false
gss2022 <- gss_get_yr(year = 2022) |>
  haven::zap_labels()

d <- gss2022 |> 
  select(cappun, polviews) |> 
  drop_na() |> 
  mutate(vcappun = if_else(cappun == 1, 
                          "Favor Capital Punishment", 
                          "Oppose Capital Punishment" ),
         vpolviews = case_when(polviews %in% c(1, 2, 3) ~ "Liberal",
                               polviews %in% c(5, 6, 7) ~ "Conservative")) |> 
  drop_na(vpolviews) |>
  select(vcappun, vpolviews)

ct <- d |>
  group_by(vcappun, vpolviews) |> 
  tally() |> 
  pivot_wider(names_from = vcappun, values_from = n) |>
  rename("Political Views" = vpolviews)

tt(ct)
```

**Marginal Probabilities:**

$P(\text{Conservative})$ = $\frac{898+274}{898+274+506+721} = 0.49$

$P(\text{Liberal})$ = $\frac{506+721}{2399} = 0.51$

$P(\text{Favor Capital Punishment})$ = $\frac{898+506}{2399} = 0.58$

$P(\text{Oppose Capital Punishment})$ = $\frac{274+721}{2399} = 0.42$

**Joint Probabilities:**

$P(\text{Conservative} \cap \text{Favor Capital Punishment})$ = $\frac{898}{2399} = 0.37$

$P(\text{Conservative} \cap \text{Oppose Capital Punishment})$ = $\frac{274}{2399} = 0.11$

$P(\text{Liberal} \cap \text{Favor Capital Punishment})$ = $\frac{506}{2399} = 0.21$

$P(\text{Liberal} \cap \text{Oppose Capital Punishment})$ = $\frac{721}{2399} = 0.30$

**Conditional Probabilities:**

$P(\text{Conservative} | \text{Favor Capital Punishment})$ = $\frac{898}{506+898} = 0.64$

$P(\text{Liberal} | \text{Favor Capital Punishment})$ = $\frac{506}{506+898} = 0.36$

$P(\text{Conservative} | \text{Oppose Capital Punishment})$ = $\frac{274}{274+721} = 0.28$

$P(\text{Liberal} | \text{Oppose Capital Punishment})$ = $\frac{721}{274+721} = 0.72$

$P(\text{Favor Capital Punishment} | \text{Conservative})$ = $\frac{898}{898+274} = 0.77$

$P(\text{Oppose Capital Punishment} | \text{Conservative})$ = $\frac{721}{274+721} = 0.23$

$P(\text{Favor Capital Punishment} | \text{Liberal})$ = $\frac{506}{506+721} = 0.41$

$P(\text{Oppose Capital Punishment} | \text{Liberal})$ = $\frac{721}{506+721} = 0.59$

# Table #2

For the second 2x2 contingency table, I pulled and recoded the variables *satfin* (respondent's financial satisfaction) and *god* (respondent's belief in a higher power) from the 2022 GSS.

```{r}
#| echo: false
gss2022 <- gss_get_yr(year = 2022) |>
  haven::zap_labels()

d2 <- gss2022 |> 
  select(god, satfin) |> 
  drop_na() |> 
  mutate(vgod = if_else(god == 3, 
                          "YES GOD", 
                          "NO GOD" ),
         vsatfin = if_else(satfin == 1, 
                           "Well Satisfied", 
                           "Not Well Satisfied")) |> 
  select(vgod, vsatfin)

ct2 <- d2 |>
  group_by(vgod, vsatfin) |> 
  tally() |> 
  pivot_wider(names_from = vgod, values_from = n) |>
  rename("Financial Satisfaction" = vsatfin)

tt(ct2)
```

**Marginal Probabilities:**

$P(\text{Not Well Satisfied})$ = $\frac{1749+311}{1749+311+590+78} = 0.76$

$P(\text{Well Satisfied})$ = $\frac{590+78}{2728} = 0.24$

$P(\text{No God})$ = $\frac{1749+590}{2728} = 0.86$

$P(\text{Yes God})$ = $\frac{311+78}{2728} = 0.14$

**Joint Probabilities:**

$P(\text{Not Well Satisfied} \cap \text{No God})$ = $\frac{1749}{2728} = 0.64$

$P(\text{Not Well Satisfied} \cap \text{Yes God})$ = $\frac{311}{2728} = 0.11$

$P(\text{Well Satisfied} \cap \text{No God})$ = $\frac{590}{2728} = 0.22$

$P(\text{Well Satisfied} \cap \text{Yes God})$ = $\frac{78}{2728} = 0.03$

**Conditional Probabilities:**

$P(\text{Not Well Satisfied} | \text{No God})$ = $\frac{1749}{590+1749} = 0.75$

$P(\text{Well Satisfied} | \text{No God})$ = $\frac{590}{590+1749} = 0.25$

$P(\text{Not Well Satisfied} | \text{Yes God})$ = $\frac{311}{311+78} = 0.80$

$P(\text{Well Satisfied} | \text{Yes God})$ = $\frac{78}{311+78} = 0.20$

$P(\text{No God} | \text{Not Well Satisfied})$ = $\frac{1749}{1749+311} = 0.85$

$P(\text{Yes God} | \text{Not Well Satisfied})$ = $\frac{311}{1749+311} = 0.15$

$P(\text{No God} | \text{Well Satisfied})$ = $\frac{590}{590+78} = 0.88$

$P(\text{Yes God} | \text{Well Satisfied})$ = $\frac{311}{1749+311} = 0.12$
